# Chapter 2: Regression and model validation

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  

```{r, echo=FALSE, message=FALSE}
students2014 <- read.csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", header = T)
#Structure of the data
str(students2014)
#Dimensions of the data
dim(students2014)
```


The data describes the relationship between learning approaches and students achievements in an introductory statistics course in Finland. The data has got 183 observation and consists of 7 variables in total, these are gender, age, attitude, deep learning, strategic learning, surface learning and points from final exam. Above you can see results from using the commands "str" and "dim".



```{r, echo=FALSE, message=FALSE}
plot(students2014)
summary(students2014$gender)
summary(students2014$age)
summary(students2014$attitude)
summary(students2014$deep)
summary(students2014$stra)
summary(students2014$surf)
summary(students2014$points)

```

Above you can see a graphical overview of the data. The summary command show that in the data there are 110 women and 56 men. The summary of the students age show that the average age is 25.51 years old. The median is 22 years. The attitude summary shows an average of 3.143, with a median of 3.2. The mean of deep, stragetic and surface learning is 3.680, 3.121 and 2.878 respectively. 

```{r, echo=FALSE, message=FALSE}
regression<-lm(points~gender+age+attitude, data=students2014)
summary(regression)
```

In this part we use a multiple regression analysis to analyse how points is affected by gender, age and attitude. The results show that only attitude have a significant statistic impact on the points from the exam. Age and gender does not play a part in the results (p-values are too high). Our model has an adjusted R-square value of 0.187, indicating that the chosen model explains 18.7 % of the variance shown in the dependent variable, points.  The residual standard error of the model is 5.315 on 162 degrees of freedom. 

Our overall model seems to be significant: F(3,162) = 13.64. 
Next, we will test the model without age and gender as explanatory variables. 

```{r, echo=FALSE, message=FALSE}
regression<-lm(points~attitude, data=students2014)
summary(regression)
```

These results show that attitude alone explains 18.56 % (adjusted R-square) of the variance in the dependent variable, points. Still our model is not perfect, but attitude is still a statistically significant variable in the model. 

```{r, echo=FALSE, message=FALSE}
my_model2 <-lm(points~gender, data=students2014)

#plot(students2014,model="Residuals vs Fitted")
#normal qqplot
#residuals vs leverage
plot(my_model2)
```
The residuals vs Fitted values shows us that our residuals are quite spread. Indicating that our data is not that good. 
The Q-Q plot shows us how the standardized residuals follow the ??

Residuals vs leverage gives us the...
